\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-11 forwards-and-backwards.tex
removed contours
fixup of first lemma

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}
In order to pass to a scaling limit, we will need to impose some 
conditions on the parameters of our model.
\begin{assumptions}%[Assumptions]
\label{def:model_setup}
We shall make the following assumptions on the parameters of our model.

@@

{\bf Reproduction parameters:}

We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
derivatives in both arguments;
\item $\gamma(x,m)$ is uniformly bounded;
\item $|\gamma(x, \smooth{\gamma} \eta(x)) \theta \int (r(y, \smooth{r} \eta(y)) - r(x, \smooth{r} \eta(x))) q_\theta(x, dy)|$
    is uniformly bounded for $x \in \IR^d$ and $\eta \in \measures$.
        \label{gamma_B_condition}
\end{enumerate}

@@

\begin{lemma}
    \label{lem:conditions on r}
If either
    \comment{which assumptions from 2.7 are in force? since in (1) we need $\gamma$ to be bounded, which is condition 3 above}
\begin{enumerate}
\item
\label{control through r} 
    $|\partial_{x_i}r(x,\smooth{r}\eta(x))|$ and 
    $|\partial_{x_ix_j}r(x,\smooth{r}\eta(x))|$
    are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
    or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
    such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
    \[
    \theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
    \]
    and
    \[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
    \leq C(\smooth{\gamma}\eta(x))^2.
    \]
\end{enumerate} 
then
$|\gamma(x, \smooth{\gamma} \eta(x)) \theta \int (r(y, \smooth{r} \eta(y)) - r(x, \smooth{r} \eta(x))) q_\theta(x, dy)|$
is uniformly bounded for $x \in \IR^d$ and $\eta \in \measures$.
\end{lemma}

@@

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.locations}
        \includegraphics{figures/ex1b/fkpp_123.locations}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \comment{remove blue curves}
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{proof}[of Lemma~\ref{lem:conditions on r}]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_1(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
Condition~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied.
We write
\begin{align*}
    r\big(y,\rho_r*\eta(y)\big) - r\big(x,\rho_r*\eta(x)\big)
    &=
    \sum_i (y_i - x_i) \partial_{x_i} \Big(r\big(x,\rho_r*\eta(x)\big)\Big) 
    \\ &\qquad {}
    +
    \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
        \partial_{x_i x_j} \Big(r\big(\smooth{r}\eta(z)\big)\Big) ,
\end{align*}
where $z = \kappa x + (1-\kappa) y$ for some $0 \le \kappa \le 1$.
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x, \rho_r*\eta(x))$ 
in this expression and so we have a bound of the form
\begin{multline*}
\Big| r\big(y,\rho_r*\eta(y)\big) - r\big(x,\rho_r*\eta(x)\big) \Big|
\\
\leq
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| 
\end{multline*}
for suitable constants $K_1$, $K_2$.
This is bounded uniformly because
$q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
\comment{typo}
u implies that $|B^\theta_1|$ is uniformly bounded;
since $\gamma$ is bounded as well, so is $|\gamma B^\theta_1|$.

Now suppose that instead
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
is satisfied.
First note that
\begin{multline*}
    r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
    \\
=
    (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
    + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
\end{multline*}
where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
So, in this case
\begin{align*}
&
\left| \theta 
    \int_{\IR^n} 
        \big\{
            r\big(x,\smooth{r}\eta(y)\big)
            -
            r\big(x,\smooth{r}\eta(x)\big)
        \big\}
    q_{\theta}(x,dy) 
\right| 
\\ &\qquad
\le
   \|r'\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))
        q_\theta(x, dy)
    \right|
    +
   \|r''\|_\infty
    \left|
        \theta \int_{\IR^d}
            (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
        q_\theta(x, dy)
    \right| .
\end{align*}
Therefore,
under Condition~\ref{control through gamma}, $|B^\theta_1(x,\eta)|$ is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$;
since $m^2\gamma(m)$ is bounded, so is $|\gamma(x,\eta) B^\theta_1(x,\eta)|$.
\end{proof}

@@

\begin{proof}
Consider the semimartingale decomposition from equation~\eqref{eqn:eta_martingale}:
\begin{align} \label{eqn:eta_f_mgale_decomp}
    \langle f, \eta^N_t \rangle
    &=
    \langle f, \eta^N_0 \rangle
    + \int_0^t \int_{\IR^d} \big\{
        \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
        + f(x) F(x, \eta^N_s)
        \big\} \eta^N_s(dx) ds
    + M^N_t(f) ,
\end{align}
where $M^N_t(f)$ is a martingale and $B^\theta_f$ is defined in \ref{generator Bf}.

We need a bound on $\gamma(x,\eta)B^\theta_f(x,\eta)+fF$.
First note that we may rewrite $B^\theta_f(x, \eta)$ as
\begin{align}
    \nonumber
&
    \left| \theta 
        \int_{\IR^n} 
            \big\{
                f(y) r\big(y,\smooth{r}\eta(y)\big)
                -
                f(x) r\big(x, \smooth{r}\eta(x)\big)
            \big\}
        q_{\theta}(x,dy)
    \right|
\\ 
    &\qquad {} =
    \Big| \theta 
        \int_{\IR^n} 
            ( f(y) - f(x) )
            r\big(y, \smooth{r}\eta(y)\big)
        q_{\theta}(x,dy)
\\ 
\nonumber
    &\qquad \qquad {}  +
    \theta 
            f(x)
        \int_{\IR^n} 
            \big\{
                r\big(y,\smooth{r}\eta(y)\big)
                -
                r\big(x,\smooth{r}\eta(x)\big)
            \big\}
        q_{\theta}(x,dy) 
    \Big| 
\\ 
    &\qquad {} \leq
    \left| \theta 
        \int_{\IR^n} 
            ( f(y) - f(x) )
            r(y, \smooth{r}\eta(y))
        q_{\theta}(x,dy)
    \right|
    +
    \|f\|_\infty |B^\theta_1(x, \eta)| .
\end{align}
\comment{TODO FIX THIS}
Using the intermediate value theorem,
then the fact that $0 \le r \le 1$,
there is a $\kappa$ in $[0, 1]$ such that the first term is
\begin{align*}
    &
    \left| \theta 
        \int_{\IR^n} 
            ( f(y) - f(x) )
            r(y, \eta)
        q_{\theta}(x,dy)
    \right|
    \\ &\qquad =
    \left| \theta 
        \int_{\IR^n} 
            \left(
                \sum_i (y-x)_i \partial_{x_i} f(x)
                +
                \frac{1}{2} \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(x + \kappa(y-x))
            \right)
            r(y, \eta)
        q_{\theta}(x,dy)
    \right|
    \\ &\qquad \leq
    \|f'\|_\infty \left| \theta 
        \int_{\IR^n} 
            (y-x) r(y, \eta)
        q_{\theta}(x,dy)
    \right|
    + \frac{1}{2} \|f''\|_\infty \left| \theta 
        \int_{\IR^n} 
            (y-x)^2
        q_{\theta}(x,dy)
    \right|
\end{align*}
bounded uniformly
because $q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$,
and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
Furthermore, Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
stipulates that $|\gamma(x, \eta) B^\theta_1(x, \eta)|$ is uniformly bounded,
and so recalling that $F$ is bounded above, we can now conclude that
under Assumptions~\ref{def:model_setup}
$\gamma(x, \eta) B^\theta_f(x, \eta) + f(x) F(x, \eta) \le C_f $ for some $C_f$.

Taking expectations in~(\ref{eqn:eta_f_mgale_decomp}),
    \begin{align}
\label{bound on intfdeta}
        \IE\left[ \langle f, \eta^N_t \rangle \right]
        &\le
        \IE\left[ \langle f, \eta^N_0 \rangle \right]
        + C_f \int_0^t \IE\left[ \langle 1, \eta^N_s \rangle \right] dt .
    \end{align}
The bound~(\ref{eta_f_bound}) then follows by first applying Gronwall's
inequality in the case $f=1$, which yields
\[
	\IE\big[\langle 1,\eta_t^N\rangle\big]\leq e^{Ct}\IE\big[\langle 1,\eta_0^N\rangle\big],
\]
with $C$ independent of $N$, and then substituting 
the resulting bound on $\IE\big[\langle 1,\eta_s^N\rangle\big]$ into the
expression above.
\end{proof}

>>>>

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}
In order to pass to a scaling limit, we will need to impose some 
conditions on the parameters of our model.

\begin{assumptions}%[Assumptions]
\label{def:model_setup}
We shall make the following assumptions on the parameters of our model.

@@

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$, there is a $C_f$ such that
    $$ |\gamma(x, \smooth{\gamma} \eta(x)) \theta \int 
    (r(y, \smooth{r} \eta(y)) f(y) - r(x, \smooth{r} \eta(x)) f(x))
    q_\theta(x, dy)| \le C_f $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
        \label{gamma_B_condition}
\end{enumerate}

@@

\begin{lemma}
    \label{lem:conditions on r}
    Assume that Conditions~\ref{def:model_setup} are satisfied,
    except for condition~\ref{gamma_B_condition}.
    If either
\begin{enumerate}
\item
\label{control through r} 
    $|\partial_{x_i}r(x,\smooth{r}\eta(x))|$ and 
    $|\partial_{x_ix_j}r(x,\smooth{r}\eta(x))|$
    are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
    or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
    such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
    \[
    \theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
    \]
    and
    \[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
    \leq C(\smooth{\gamma}\eta(x))^2 ,
    \]
\end{enumerate} 
    then condition~\ref{gamma_B_condition} is also satisfied.
\end{lemma}

@@

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        \includegraphics{figures/ex1b/fkpp_123.density.3.0}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

@@

\begin{proof}[of Lemma~\ref{lem:conditions on r}]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
Condition~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied.
We write
\begin{align*}
    r\big(y,\rho_r*\eta(y)\big) f(y) - r\big(x,\rho_r*\eta(x)\big) f(x)
    &=
    \sum_i (y_i - x_i) \partial_{x_i} \Big( r\big(x,\rho_r*\eta(x)\big) f(x) \Big) 
    \\ &\qquad {}
    +
    \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
        \partial_{x_i x_j} \Big( r\big(z, \smooth{r}\eta(z)\big) f(z) \Big) ,
\end{align*}
where $z = \kappa x + (1-\kappa) y$ for some $0 \le \kappa \le 1$.
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x, \rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{multline*}
|B^\theta_f|
\leq
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| 
\end{multline*}
for suitable constants $K_1$, $K_2$.
This is bounded uniformly because
$q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
Since $\gamma$ is bounded as well, so is $|\gamma B^\theta_f|$.

Now suppose that instead
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
First note that
    \begin{align}
	    \nonumber
    & |B_f^\theta| =
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x, \smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
 \label{eqn:q_bound_split}
	    &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y, \smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
\nonumber
	    &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

 Writing
    $K_3 = \sup_{x,m} \max_i |\partial_{x_i} f(x)r(x, m)|$
    and 
    $K_4 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x)r(x, m)|$, since
\[\big|\partial_{x_i}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_3\]
and
\[\big|\partial_{x_ix_j}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_4,\]
the first term
is bounded exactly as above. For the second, 
    \begin{multline*}
        r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
        \\
=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
    \end{multline*}
    where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
    we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
    So, writing $K_5=\|f\|_\infty\|r'\|_\infty$ and $K_6=\|f\|_\infty\|r''\|_\infty$,
    the second term in~\eqref{eqn:q_bound_split} is bounded by
    \begin{align*}
       K_5
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
       K_6
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
Under Condition~\ref{control through gamma} of 
Assumptions~\ref{def:model_setup}, the right hand side is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$ and 
$m^2\gamma(m)$ is bounded, so therefore $|\gamma B^\theta_f|$ is bounded.
\end{proof}

@@

\begin{proof}
Consider the semimartingale decomposition from equation~\eqref{eqn:eta_martingale}:
\begin{align} \label{eqn:eta_f_mgale_decomp}
    \langle f, \eta^N_t \rangle
    &=
    \langle f, \eta^N_0 \rangle
    + \int_0^t \int_{\IR^d} \big\{
        \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
        + f(x) F(x, \eta^N_s)
        \big\} \eta^N_s(dx) ds
    + M^N_t(f) ,
\end{align}
where $M^N_t(f)$ is a martingale and $B^\theta_f$ is defined in \ref{generator Bf}.
First note that Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
stipulates that $|\gamma B^\theta_f|$ is uniformly bounded,
and so recalling that $F$ is bounded above, we can now conclude that
under Assumptions~\ref{def:model_setup}
$\gamma(x, \eta) B^\theta_f(x, \eta) + f(x) F(x, \eta) \le C_f $ for some $C_f$.

Now, taking expectations in~(\ref{eqn:eta_f_mgale_decomp}),
    \begin{align}
\label{bound on intfdeta}
        \IE\left[ \langle f, \eta^N_t \rangle \right]
        &\le
        \IE\left[ \langle f, \eta^N_0 \rangle \right]
        + C_f \int_0^t \IE\left[ \langle 1, \eta^N_s \rangle \right] dt .
    \end{align}
The bound~(\ref{eta_f_bound}) then follows by first applying Gronwall's
inequality in the case $f=1$, which yields
\[
	\IE\big[\langle 1,\eta_t^N\rangle\big]\leq e^{Ct}\IE\big[\langle 1,\eta_0^N\rangle\big],
\]
with $C$ independent of $N$, and then substituting 
the resulting bound on $\IE\big[\langle 1,\eta_s^N\rangle\big]$ into the
expression above.
\end{proof}
